Clustering in a hyperbolic model of complex networks.

### 2020-01-02 concentration_argument.tex
version 1.6

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$ and $k_n \to \infty$ such that $k_n = \smallO{n}$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and define
\[
	\lambda_n^\pm = (\ell_n \pm C \sqrt{\ell_n \log(\ell_n)}) \wedge \xi, \quad \text{and} \quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right).
\] 
In addition, let $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, where $\hat{\mu}_n(y)$ is a differentiable function that satisfies for some $0 < \varepsilon < 1$
\begin{align*}
	&\mathrm{i)} \, \lim_{n \to \infty} \sup_{0 \le y \le (1 - \varepsilon)R} \frac{\hat{\mu}_n(y)}{\mu(\BallPo{y})} = 1
	\quad \text{and}
	&&\mathrm{ii)} \, \lim_{n \to \infty}  \sup_{0 \le y \le (1 - \varepsilon)R} \frac{\hat{\mu}_n^\prime(y)}{\mu(\BallPo{y})} = \frac{1}{2}.
\end{align*}
%
%
% with $\hat{\mu}_n(y) = (1 + \phi_n(y))\mu(y)$ and 
%where $\phi_n(y)$ is a differentiable function satisfying:
%\begin{enumerate}[\upshape i)]
%\item For some $0 < \varepsilon < 1$, $\displaystyle \lim_{n \to \infty} \sup_{0 \le y \le (1 - \varepsilon)R} |\phi_n(y)| = 0$,
%\item $\displaystyle \lim_{n \to \infty}  \sup_{y \in \Kcal_{C}(\ell_n)} |\phi_n^\prime(y)| = 0.$
%\end{enumerate}
Then the following holds:
\begin{enumerate}
\item For any continuous function $h : \R_+ \rightarrow  \R$, such that $h(y) = \bigO{e^{\beta y}}$ as $y \to \infty$ for some $\beta < \alpha$, 
\begin{equation}\label{eq:error_bound_int_rho_approx_not_K}
	\int_{\R_+ \setminus (a_n^-, a_n^+)} \hat{\rho}_n(y,k_n) h(y) \alpha e^{-\alpha y} \dd y
	= \bigO{k_n^{-(1+C^2)/2}},
\end{equation}
as $n \to \infty$.
\item If in addition, $h(a_n) \sim h(b_n)$ whenever $a_n \sim b_n$, as $n \to \infty$, then,
\begin{equation}\label{eq:concentration_h_rho_approx}
	\int_0^\infty h(y) \hat{\rho}_n(y,k_n) \alpha e^{-\alpha y} \dd y \sim  
		\int_0^\infty h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y,
\end{equation}
as $n \to \infty$.
\end{enumerate}
\end{lemma}

@@

For simplicity we write $\mu(y) := \mu\left(\BallPo{y}\right) = \xi e^{y/2}$ throughout the proof. Observe that $\mu^\prime(y) = \mu(y)/2$ and $\mu^{-1}(y z) = \mu^{-1(y)} + \mu^{-1}(z)$.

%We note that the statement will follow from Lemma~\ref{lem:concentration_argument} if we can show that
%\[
%	\int_0^R \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
%	= \left(1 + o(1)\right) \int_0^R \rho(z,k_n) e^{-\alpha z} \dd z.
%\] 
%or equivalently,
%\[
%	\int_0^R \rho(z,k_n) e^{-\alpha z} \dd z
%	= \left(1 + o(1)\right)\int_0^R \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y.
%\]
%
%We will do this in two steps.
%
%Therefore, what is left is to show that
%\[
%	\int_0^{(1-\varepsilon^\prime)R} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
%	= \left(1 + o(1)\right) \int_0^{(1-\varepsilon^\prime)R} \rho(z,k_n) e^{-\alpha z} \dd z.
%\] 
%or equivalently,
%\[
%	\int_0^{(1-\varepsilon^\prime)R} \rho(z,k_n) e^{-\alpha z} \dd z
%	= \left(1 + o(1)\right)\int_0^{(1-\varepsilon^\prime)R} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y.
%\]
%
%Using the change of variable $z = \mu^{-1}(\hat{\mu}_n(y))$ and writing $\hat{a}_n^\pm = \hat{\mu}_n^{-1}(\mu(a_n^\pm))$ and $\hat{R}_{\varepsilon^\prime} := \hat{\mu}_n^{-1}(\mu((1-\varepsilon^\prime)R))$, we get
%\begin{align*}
%	\int_{a_n^+}^{(1-\varepsilon^\prime)R} \rho(z,k_n) e^{-\alpha z} \dd z 
%	&= \int_{a_n^+}^{(1-\varepsilon^\prime)R} \Prob{\Po(\mu(z)) = k_n} e^{-\alpha z} \dd z\\
%	&= \int_{\hat{a}_n^+}^{\hat{R}_{\varepsilon^\prime}} \Prob{\Po(\hat{\mu}_n(y)) = k_n} e^{-\alpha \mu^{-1}(\hat{\mu}_n(y))} 
%		\frac{\hat{\mu}_n^\prime(y)}{\mu^\prime(\mu^{-1}(\hat{\mu}_n(y)))} \dd y,
%\end{align*}
%where the fraction in the last line follows from the chain rule and the fact that $(\mu^{-1})^\prime(t) = (\mu^\prime(\mu^{-1}(t)))^{-1}$.

@@

Let $0 < \varepsilon < 1$ be such that conditions i) and ii) hold, take $\varepsilon^\prime = \min\{\varepsilon/3,1/3\} < 1/2$ and let $\delta > 0$ be such that $\hat{\mu}_n(\delta) \ge \xi$. We first show that we can restrict to integration over $(\delta, (1-\varepsilon^\prime)R)$ that the integration over $(1-\varepsilon^\prime)R \le y \le R$ is negligible.

Since $h(y) = \bigO{e^{\beta y}}$ we have
\begin{align*}
	\int_{(1-\varepsilon^\prime)R}^{R} h(y) \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	&= \bigO{1} \hat{\rho}_n((1-\varepsilon^\prime)R,k_n) e^{-(\alpha-\beta)(1-\varepsilon^\prime)R} \\
	&= \bigO{\hat{\rho}_n((1-\varepsilon^\prime)R,k_n) n^{-2(\alpha-\beta)(1-\varepsilon^\prime)}}.
\end{align*}
By construction $\varepsilon^\prime < \varepsilon$, and hence by condition i) we have that $\hat{\mu}_n((1-\varepsilon^\prime)R) = \bigT{\mu((1-\varepsilon^\prime)R} = \bigT{n^{2(1-\varepsilon^\prime)}}$. Therefore, since $k_n = \smallO{n}$, it follows that $\hat{\mu}_n((1-\varepsilon^\prime)R)/k_n = \omega\left(n^{1 - 2\varepsilon^\prime}\right) \to \infty$ as $n \to \infty$. We can now use Stirling's approximation to bound $\hat{\rho}_n((1-\varepsilon^\prime)R,k_n)$ as
\begin{align*}
	\hat{\rho}_n((1-\varepsilon^\prime)R,k_n) &= \Prob{\Po(\hat{\mu}_n((1-\varepsilon^\prime)R)) = k_n} \\
	&= \frac{\hat{\mu}_n((1-\varepsilon^\prime)R)^{k_n}}{k_n!} e^{-\hat{\mu}_n((1-\varepsilon^\prime)R)}\\
	&= \bigO{1} k_n^{-1/2} \left(\frac{\hat{\mu}_n((1-\varepsilon^\prime)R)}{k_n}\right)^{k_n} e^{k_n - \hat{\mu}_n((1-\varepsilon^\prime)R)}\\
	&= \bigO{1} k_n^{-1/2} e^{k_n\left(1 - \frac{\hat{\mu}_n((1-\varepsilon^\prime)R)}{k_n}\right)+ \log\left(\frac{\hat{\mu}_n((1-\varepsilon^\prime)R)}{k_n}\right)}\\
	&\le \bigO{1} k_n^{-1/2} e^{-\hat{\mu}_n((1-\varepsilon^\prime)R)/2},
\end{align*}
where the last line follows since $1 - x + \log(x) \le -x/2$ for large enough $x$. We conclude that
\[
	\int_{(1-\varepsilon^\prime)R}^{R} h(y) \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	= \bigO{k_n^{-1/2} n^{-2(\beta-\alpha)(1-\varepsilon^\prime)} e^{-n^{2(1-\varepsilon^\prime)}}}
	= \bigO{k_n^{-(1+C^2)/2}}.
\]

In a similar fashion we get that
\begin{align*}
	\int_0^\delta h(y) \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	&= \bigO{1} \hat{\rho}_n(\delta, k_n) \\
	&= \bigO{1} k_n^{-1/2} e^{-k_n(\log(k_n)-\hat{\mu}_n(\delta))} = \bigO{k_n^{-(1+C^2)/2}}.
\end{align*}

@@

Let $(a, b) \subseteq (\delta,(1-\varepsilon^\prime)R)$ and consider the change of variable $z = \mu^{-1}(\hat{\mu}_n(y))$. Then, writing $\hat{a} = \hat{\mu}_n^{-1}(\mu(a))$ and similar for $\hat{b}$, we get
\begin{align*}
	\int_{a}^{b} \rho(z,k_n) e^{-\alpha z} \dd z 
	&= \int_{a}^{b} \Prob{\Po(\mu(z)) = k_n} e^{-\alpha z} \dd z\\
	&= \int_{\hat{a}}^{\hat{b}} \Prob{\Po(\hat{\mu}_n(y)) = k_n} e^{-\alpha \mu^{-1}(\hat{\mu}_n(y))} 
		\frac{\hat{\mu}_n^\prime(y)}{\mu^\prime(\mu^{-1}(\hat{\mu}_n(y)))} \dd y,
\end{align*}
where the fraction in the last line follows from the chain rule and the fact that $(\mu^{-1})^\prime(t) = (\mu^\prime(\mu^{-1}(t)))^{-1}$.

@@

Now, using~\eqref{eq:equivalence_integral_rho_change_variables} with $a = \mu^{-1}(\hat{\mu}_n(\delta))$ and $b = \mu^{-1}(\hat{\mu}_n(a_n^-)$ we get
\begin{align*}
	\int_a^b \rho(y,k_n) e^{(\beta-\alpha) y} \dd y
	&= (1+\smallO{1}) \int_{\delta}^{a_n^-} \hat{\rho}_n(y,k_n) e^{(\beta - \alpha)y} \dd y.
\end{align*}
Since $\mu^{-1}(\hat{\mu}_n(a_n^-) = (1+\smallO{1})a_n^-$ and $b \ge 0$ the left hand side is
\[
	\bigO{1} \int_0^{a_n^-} \rho(y,k_n) e^{(\beta-\alpha) y} \dd y,
\]
and hence~\eqref{eq:concentration_argument_an_-} follows from Lemma~\ref{lem:concentration_argument}. The proof of~\eqref{eq:concentration_argument_an_+} follows in a similar way.

@@

The proof of the second statement follows the same line of reasoning as above, using~\eqref{eq:equivalence_integral_rho_change_variables}. First, the mean value theorem for definite integrals implies that
\[
	\int_{a_n^-}^{a_n^+} \hat{\rho}_n(y,k_n) h(y) e^{-\alpha y} \dd y
	= h(c_n) \int_{a_n^-}^{a_n^+} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y,
\]
for some $c_n \in (a_n^-, a_n^+)$. Since, $c_n \sim 2 \log(k_n/\xi)$, by assumption on $h$, $h(c_n)\sim h(2\log(k_n/\xi))$. 
Therefore it is enough to show that
\[
	\int_{a_n^-}^{a_n^+} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	= \left(1 + o(1)\right) \int_{a_n^-}^{a_n^+} \rho(z,k_n) e^{-\alpha z} \dd z.
\] 
This however follows immediately from~\eqref{eq:equivalence_integral_rho_change_variables} by picking $a = \mu^{-1}(\hat{\mu}_n(a_n^-))$ and $b = \mu^{-1}(\hat{\mu}_n(a_n^+))$.
\end{proof}

@@

A direct consequence of Lemma~\ref{lem:average_degree_hyperbolic} is that $\Mu{\BallHyp{y}} = \Mu{\BallPo{y}}(1 + \phi_n(y))$, where $\phi_n(y) := \Mu{\BallHyp{y}}/\Mu{\BallPo{y}} - 1$ satisfies condition i) in Lemma~\ref{lem:concentration_argument_rho_approximation}. To show that condition ii) is also satisfied we need to analyze
\[
	\phi_n^\prime(y) = 
	\Mu{\BallPo{y}}^{-1} \frac{\partial}{\partial y} \mu\left(\BallHyp{y}\right) -  \frac{1}{2} \frac{\Mu{\BallHyp{y}}}{\Mu{\BallPo{y}}},
\]
where we used that $\frac{\partial}{\partial y}\Mu{\BallPo{y}} = \frac{1}{2}\Mu{\BallPo{y}}$. Again, Lemma~\ref{lem:average_degree_hyperbolic} implies that 
\[
	\lim_{n \to \infty} \sup_{0 \le y \le (1-\varepsilon)R} \frac{1}{2} \frac{\Mu{\BallHyp{y}}}{\Mu{\BallPo{y}}}
	= \frac{1}{2}.
\]
The following lemma shows that the same holds for the first term from which we conclude that $\phi_n(y)$ satisfies condition ii) in Lemma~\ref{lem:concentration_argument_rho_approximation}.

\begin{lemma}
For any $0 < \varepsilon < 1$,
\[
	\lim_{n \to \infty} \sup_{0 \le y \le (1-\varepsilon)R} \Mu{\BallPo{y}}^{-1}
	\frac{\partial}{\partial y} \mu\left(\BallHyp{y}\right) 
	= \frac{1}{2}.
\]
\end{lemma}

\begin{proof}
We again split $\Mu{\BallHyp{y}}$ over the top and bottom part,
\[
	\Mu{\BallHyp{y}} 
	= \Mu{\BallHyp{y} \cap \Rcal ([0,R - y))} + \Mu{\BallHyp{y} \cap \Rcal ([R - y, R])},
\]
where
\[
	\Mu{\BallHyp{y} \cap \Rcal ([0,R - y))} = \frac{2\alpha \nu}{\pi}\int_0^{R - y} \Phi(y,y^\prime) 
		e^{-\alpha y^\prime} \dd y^\prime,
\]
with $\Phi(y,y^\prime)$ defined as in~\eqref{eq:def_Omega_hyperbolic} and, see~\eqref{eq:mu_hyperbolic_ball_part_1},
\[
	\Mu{\BallHyp{y} \cap \Rcal ([R - y, R])}
	= \xi e^{y/2}\frac{2\alpha - 1}{4\pi} \left( e^{-(\alpha - \frac{1}{2})(R - y)}
	- e^{-(\alpha - \frac{1}{2})R - y/2}\right).
\]
Taking the derivative of the last expression gives
\begin{align*}
	&\frac{\partial}{\partial y} \Mu{\BallHyp{y} \cap \Rcal ([R - y, R])}\\
	&= \frac{1}{2}\Mu{\BallHyp{y} \cap \Rcal ([R - y, R])}
		+ \xi e^{y/2}\frac{2\alpha - 1}{4\pi}\left(
		\left(\alpha - \frac{1}{2}\right)e^{-(\alpha - \frac{1}{2})(R - y)} 
		+ \frac{1}{2}e^{-(\alpha - \frac{1}{2})R - y/2}\right)\\
	&= \frac{1}{2}\Mu{\BallHyp{y} \cap \Rcal ([R - y, R])}\left(1 +
		\frac{(2\alpha - 1)e^{-(\alpha - \frac{1}{2})(R - y)} + e^{-(\alpha - \frac{1}{2})R - y/2}}
		{e^{-(\alpha - \frac{1}{2})(R - y)} - e^{-(\alpha - \frac{1}{2})R - y/2}}\right).
\end{align*}
Since, $\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R} \Mu{\BallPo{y}}^{-1} \Mu{\BallHyp{y} \cap \Rcal ([R - y, R])} = 0$, we are left to show that
\begin{equation}\label{eq:derivative_mu_hyp_ball_main}
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R} \Mu{\BallPo{y}}^{-1} \frac{2\alpha \nu}{\pi} \frac{\partial}{\partial y} \int_0^{R - y} \Phi(r,r^\prime) e^{-\alpha y^\prime} \dd y^\prime
	= \frac{1}{2}.
\end{equation}

We start with some preliminary computations. For convenience we define
\[
	\Xi(y,y^\prime) = 1 - \frac{\cosh(R- y)\cosh(R-y^\prime) - \cosh(R)}{\sinh(R - y) \sinh(R - y^\prime)},
\]
so that
\[
	\Phi(R -y, R - y^\prime) = \frac{1}{2}e^{R/2} \arccos\left(1 - \Xi(y,y^\prime)\right).
\]
Next, following the same calculation as in the proof of~\cite[Lemma 28]{fountoulakis2018law}, we write
\begin{align*}
	\Xi(y,y^\prime)
	&= 2 e^{-(R - y - y^\prime)} \frac{\left(1 - e^{y^\prime - y - R}\right)\left(1 - e^{y - y^\prime - R}\right)}
		{\left(1 - e^{-2(R - y^\prime)}\right)\left(1 - e^{-2(R- y)}\right)}\\
	&:= 2 e^{-(R - y - y^\prime)} \frac{h_1(y) h_2(y)}{h_3(y^\prime) h_3(y)},
\end{align*}
with
\[
	h_1(y) = 1 - e^{y^\prime - y - R}, \quad h_2(y) = 1 - e^{y - y^\prime - R}
	\quad \text{and} \quad h_3(y) = 1 - e^{-2(R- y)}.
\]
We suppressed the dependence on $n$ and, in some cases, on $y^\prime$ for notation convenience.

We make two important observations. First, $\Xi(y,y^\prime)$ is an increasing function in both arguments, for $y, y^\prime < R$ and $y + y^\prime < R$. Second, for all $y + y^\prime < R$, $h_1(y) \le h_3(y^\prime)$ and $h_2(y) \le h_3(y)$, while $h_3(y), h_3(y^\prime) < 1$, so that
\begin{equation}\label{eq:derivative_hyp_ball_Xi_bounds}
	2 e^{-(R - y - y^\prime)}h_1(y) h_2(y) \le \Xi(y,y^\prime) \le 2 e^{-(R - y - y^\prime)}.
\end{equation}
In particular, there exists a $0 < \delta < 1$ such that $1 \le \Xi(y,y^\prime) \le 2$ for all $0 < y < R$ and $(1-\delta)(R-y) < y^\prime < R$.

>>>>

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$ and $0 < \varepsilon < 1$. Furthermore, let $k_n \to \infty$ such that $k_n = \bigO{n^{1-\varepsilon}}$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and define
\[
	\lambda_n^\pm = (\ell_n \pm C \sqrt{\ell_n \log(\ell_n)}) \wedge \xi, \quad \text{and} \quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right),
\] 
for some $C > 0$. Finally, let $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, where $\hat{\mu}_n(y)$ is a differentiable function that satisfies for some $0 < \varepsilon^\prime < 1$
\begin{align*}
	&\mathrm{i)} \, \lim_{n \to \infty} \sup_{0 \le y \le (1 - \varepsilon^\prime)R} 
		\left|\frac{\hat{\mu}_n(y)}{\mu(\BallPo{y})} - 1\right| = 0
	\quad \text{and}
	&&\mathrm{ii)} \, \lim_{n \to \infty}  \sup_{0 \le y \le (1 - \varepsilon^\prime)R} 	
		\left|\frac{\hat{\mu}_n^\prime(y)}{\mu(\BallPo{y})} - \frac{1}{2}\right| = 0.
\end{align*}
Then the following holds for $C > 0$ large enough:
\begin{enumerate}
\item For any continuous function $h : \R_+ \rightarrow  \R$, such that $h(y) = \bigO{e^{\beta y}}$ as $y \to \infty$ for some $\beta < \alpha$, 
\begin{equation}\label{eq:error_bound_int_rho_approx_not_K}
	\int_{\R_+ \setminus (a_n^-, a_n^+)} \hat{\rho}_n(y,k_n) h(y) \alpha e^{-\alpha y} \dd y
	= \bigO{k_n^{-(1+C^2)/2}},
\end{equation}
as $n \to \infty$.
\item If in addition, $h(a_n) \sim h(b_n)$ whenever $a_n \sim b_n$, as $n \to \infty$, then,
\begin{equation}\label{eq:concentration_h_rho_approx}
	\int_0^\infty h(y) \hat{\rho}_n(y,k_n) \alpha e^{-\alpha y} \dd y \sim  
		\int_0^\infty h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y,
\end{equation}
as $n \to \infty$.
\end{enumerate}
\end{lemma}

@@

For simplicity we write $\mu(y) := \mu\left(\BallPo{y}\right) = \xi e^{y/2}$ throughout the proof. Observe that $\mu^\prime(y) = \mu(y)/2$ and $\mu^{-1}(y z) = \mu^{-1}(y) + \mu^{-1}(z)$.

@@

Let $0 < \varepsilon < 1$ be such that conditions i) and ii) hold, 
Take any $0 < \delta < \min\{\varepsilon, \varepsilon^\prime/3,1/3\} < 1/2$ and let $Q > 0$ be such that $\hat{\mu}_n(Q) \ge \xi$. We first show that we can restrict to integration over $(Q, (1-\delta)R)$, starting with showing that the integration over $(1-\delta)R \le y \le R$ is negligible.

By construction $\delta < \varepsilon^\prime$, and hence by condition i) we have that $\hat{\mu}_n((1-\delta)R) = \bigT{\mu((1-\delta)R} = \bigT{n^{(1-\delta)}}$. Therefore, since $\delta<\varepsilon$ and $k_n = \bigO{n^{1-\varepsilon}}$, it follows that $\hat{\mu}_n((1-\delta)R)/k_n = \omega\left(n^{\varepsilon - \delta}\right)$ as $n \to \infty$. In particular $\hat{\mu}((1-\delta)R) = \omega{k_n}$ and hence $\hat{\rho}_n(y,k_n) \le \hat{\rho}((1-\delta)R,k_n)$ for all $y \ge (1-\delta)R$. It now follow that
\begin{align*}
	\int_{(1-\delta)R}^{R} h(y) \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	&= \bigO{1} \hat{\rho}_n((1-\delta)R,k_n) e^{-(\alpha-\beta)(1-\delta)R} \\
	&= \bigO{\hat{\rho}_n((1-\delta)R,k_n) n^{-2(\alpha-\beta)(1-\delta)}},
\end{align*}
where we used that that $h(y) = \bigO{e^{\beta y}}$.
Using Stirling's approximation to bound $\hat{\rho}_n((1-\delta)R,k_n)$ we get
\begin{align*}
	\hat{\rho}_n((1-\delta)R,k_n) &= \Prob{\Po(\hat{\mu}_n((1-\delta)R)) = k_n} \\
	&= \frac{\hat{\mu}_n((1-\delta)R)^{k_n}}{k_n!} e^{-\hat{\mu}_n((1-\delta)R)}\\
	&= \bigO{1} k_n^{-1/2} \left(\frac{\hat{\mu}_n((1-\delta)R)}{k_n}\right)^{k_n} e^{k_n - \hat{\mu}_n((1-\delta)R)}\\
	&= \bigO{1} k_n^{-1/2} e^{k_n\left(1 - \frac{\hat{\mu}_n((1-\delta)R)}{k_n}+ \log\left(\frac{\hat{\mu}_n((1-\delta)R)}{k_n}\right)\right)}\\
	&\le \bigO{1} k_n^{-1/2} e^{-\hat{\mu}_n((1-\delta)R)/2},
\end{align*}
where the last line follows since $1 - x + \log(x) \le -x/2$ for large enough $x$ and $\hat{\mu}_n((1-\delta)R)/k_n \to \infty$. We thus conclude that for any $C > 0$
\[
	\int_{(1-\delta)R}^{R} h(y) \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	= \bigO{k_n^{-1/2} n^{-2(\beta-\alpha)(1-\delta)} e^{-n^{(1-\delta)}}}
	= \bigO{k_n^{-(1+C^2)/2}}.
\]

For the range $(0,Q)$ we have, for any $C > 0$,
\begin{align*}
	\int_0^Q h(y) \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	&= \bigO{1} \hat{\rho}_n(Q, k_n) \\
	&= \bigO{1} k_n^{-1/2} e^{-k_n(\log(k_n)-\hat{\mu}_n(Q))} = \bigO{k_n^{-(1+C^2)/2}}.
\end{align*}

@@

Let $(a, b) \subseteq (Q,(1-\delta)R)$ and consider the change of variable $z = \mu^{-1}(\hat{\mu}_n(y))$. Then, writing $\hat{a} = \hat{\mu}_n^{-1}(\mu(a))$ and similar for $\hat{b}$, we get
\begin{align*}
	\int_{a}^{b} \rho(z,k_n) e^{-\alpha z} \dd z 
	&= \int_{a}^{b} \Prob{\Po(\mu(z)) = k_n} e^{-\alpha z} \dd z\\
	&= \int_{\hat{a}}^{\hat{b}} \Prob{\Po(\hat{\mu}_n(y)) = k_n} e^{-\alpha \mu^{-1}(\hat{\mu}_n(y))} 
		\frac{\hat{\mu}_n^\prime(y)}{\mu^\prime(\mu^{-1}(\hat{\mu}_n(y)))} \dd y,
\end{align*}
where the fraction in the last line follows from the chain rule and the fact that $(\mu^{-1})^\prime(t) = (\mu^\prime(\mu^{-1}(t)))^{-1}$.

@@

Now, using~\eqref{eq:equivalence_integral_rho_change_variables} with $a = \mu^{-1}(\hat{\mu}_n(Q))$ and $b = \mu^{-1}(\hat{\mu}_n(a_n^-)$ we get
\begin{align*}
	\int_a^b \rho(y,k_n) e^{(\beta-\alpha) y} \dd y
	&= (1+\smallO{1}) \int_{Q}^{a_n^-} \hat{\rho}_n(y,k_n) e^{(\beta - \alpha)y} \dd y.
\end{align*}
Since $\mu^{-1}(\hat{\mu}_n(a_n^-) = (1+\smallO{1})a_n^-$ and $b \ge 0$ the left hand side is
\[
	\bigO{1} \int_0^{a_n^-} \rho(y,k_n) e^{(\beta-\alpha) y} \dd y,
\]
and hence~\eqref{eq:concentration_argument_an_-} follows from Lemma~\ref{lem:concentration_argument}. The proof of~\eqref{eq:concentration_argument_an_+} follows in a similar way.

@@

The proof of the second statement follows the same line of reasoning as above, using~\eqref{eq:equivalence_integral_rho_change_variables}. First, the mean value theorem for definite integrals implies that
\[
	\int_{a_n^-}^{a_n^+} \hat{\rho}_n(y,k_n) h(y) e^{-\alpha y} \dd y
	= h(c_n) \int_{a_n^-}^{a_n^+} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y,
\]
for some $c_n \in (a_n^-, a_n^+)$. Since $c_n \sim 2 \log(k_n/\xi)$, $h(c_n)\sim h(2\log(k_n/\xi))$ by assumption on $h$. 
Therefore it is enough to show that
\[
	\int_{a_n^-}^{a_n^+} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	= \left(1 + o(1)\right) \int_{a_n^-}^{a_n^+} \rho(z,k_n) e^{-\alpha z} \dd z.
\] 
This however follows immediately from~\eqref{eq:equivalence_integral_rho_change_variables} by picking $a = \mu^{-1}(\hat{\mu}_n(a_n^-))$ and $b = \mu^{-1}(\hat{\mu}_n(a_n^+))$.
\end{proof}

@@

A direct consequence of Lemma~\ref{lem:average_degree_hyperbolic} is that $\Mu{\BallHyp{y}} = \Mu{\BallPo{y}}(1 + \phi_n(y))$, where $\phi_n(y) := \Mu{\BallHyp{y}}/\Mu{\BallPo{y}} - 1$ satisfies condition i) in Lemma~\ref{lem:concentration_argument_rho_approximation}. To show that condition ii) is also satisfied we need to analyze
\[
	\phi_n^\prime(y) = 
	\Mu{\BallPo{y}}^{-1} \frac{\partial}{\partial y} \mu\left(\BallHyp{y}\right) -  \frac{1}{2} \frac{\Mu{\BallHyp{y}}}{\Mu{\BallPo{y}}},
\]
where we used that $\frac{\partial}{\partial y}\Mu{\BallPo{y}} = \frac{1}{2}\Mu{\BallPo{y}}$. Again, Lemma~\ref{lem:average_degree_hyperbolic} implies that 
\[
	\lim_{n \to \infty} \sup_{0 \le y \le (1-\varepsilon)R} \left|\frac{1}{2} \frac{\Mu{\BallHyp{y}}}{\Mu{\BallPo{y}}}
	- \frac{1}{2}\right| = 0.
\]
The following lemma shows that the same holds for the first term from which we conclude that $\phi_n(y)$ satisfies condition ii) in Lemma~\ref{lem:concentration_argument_rho_approximation}.

\begin{lemma}
For any $0 < \varepsilon < 1$,
\[
	\lim_{n \to \infty} \sup_{0 \le y \le (1-\varepsilon)R} \left|\Mu{\BallPo{y}}^{-1}
	\frac{\partial}{\partial y} \mu\left(\BallHyp{y}\right) - \frac{1}{2}\right| = 0.
\]
\end{lemma}

\begin{proof}
We again split $\Mu{\BallHyp{y}}$ over the top and bottom part,
\[
	\Mu{\BallHyp{y}} 
	= \Mu{\BallHyp{y} \cap \Rcal ([0,R - y))} + \Mu{\BallHyp{y} \cap \Rcal ([R - y, R])},
\]
where
\[
	\Mu{\BallHyp{y} \cap \Rcal ([0,R - y))} = \frac{2\alpha \nu}{\pi}\int_0^{R - y} \Phi(y,y^\prime) 
		e^{-\alpha y^\prime} \dd y^\prime,
\]
with $\Phi(y,y^\prime)$ defined as in~\eqref{eq:def_Omega_hyperbolic} and, see~\eqref{eq:mu_hyperbolic_ball_part_1},
\[
	\Mu{\BallHyp{y} \cap \Rcal ([R - y, R])}
	= \xi e^{y/2}\frac{2\alpha - 1}{4\pi} \left( e^{-(\alpha - \frac{1}{2})(R - y)}
	- e^{-(\alpha - \frac{1}{2})R - y/2}\right).
\]
Taking the derivative of the last expression gives
\begin{align*}
	&\frac{\partial}{\partial y} \Mu{\BallHyp{y} \cap \Rcal ([R - y, R])}\\
	&= \frac{1}{2}\Mu{\BallHyp{y} \cap \Rcal ([R - y, R])}
		+ \xi e^{y/2}\frac{2\alpha - 1}{4\pi}\left(
		\left(\alpha - \frac{1}{2}\right)e^{-(\alpha - \frac{1}{2})(R - y)} 
		+ \frac{1}{2}e^{-(\alpha - \frac{1}{2})R - y/2}\right)\\
	&= \frac{1}{2}\Mu{\BallHyp{y} \cap \Rcal ([R - y, R])}\left(1 +
		\frac{(2\alpha - 1)e^{-(\alpha - \frac{1}{2})(R - y)} + e^{-(\alpha - \frac{1}{2})R - y/2}}
		{e^{-(\alpha - \frac{1}{2})(R - y)} - e^{-(\alpha - \frac{1}{2})R - y/2}}\right).
\end{align*}
Since, $\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R} \Mu{\BallPo{y}}^{-1} \Mu{\BallHyp{y} \cap \Rcal ([R - y, R])} = 0$, we are left to show that
\begin{equation}\label{eq:derivative_mu_hyp_ball_main}
	\lim_{n \to \infty} \sup_{0 < y \le (1-\varepsilon)R} \left|\Mu{\BallPo{y}}^{-1} \frac{2\alpha \nu}{\pi} \frac{\partial}{\partial y} \int_0^{R - y} \Phi(y,y^\prime) e^{-\alpha y^\prime} \dd y^\prime
	- \frac{1}{2}\right| = 0.
\end{equation}

We start with some preliminary computations. For convenience we define
\[
	\Xi(y,y^\prime) = 1 - \frac{\cosh(R- y)\cosh(R-y^\prime) - \cosh(R)}{\sinh(R - y) \sinh(R - y^\prime)},
\]
so that
\[
	\Phi(y, y^\prime) = \frac{1}{2}e^{R/2} \arccos\left(1 - \Xi(y,y^\prime)\right).
\]
Next, following the same calculation as in the proof of~\cite[Lemma 28]{fountoulakis2018law}, we write
\begin{align*}
	\Xi(y,y^\prime)
	&= 2 e^{-(R - y - y^\prime)} \frac{\left(1 - e^{y^\prime - y - R}\right)\left(1 - e^{y - y^\prime - R}\right)}
		{\left(1 - e^{-2(R - y^\prime)}\right)\left(1 - e^{-2(R- y)}\right)}\\
	&:= 2 e^{-(R - y - y^\prime)} \frac{h_1(y) h_2(y)}{h_3(y^\prime) h_3(y)},
\end{align*}
with
\[
	h_1(y) = 1 - e^{y^\prime - y - R}, \quad h_2(y) = 1 - e^{y - y^\prime - R}
	\quad \text{and} \quad h_3(y) = 1 - e^{-2(R- y)}.
\]
We suppressed the dependence on $n$ and, in some cases, on $y^\prime$ for notation convenience.

We make two important observations. First, $\Xi(y,y^\prime)$ is an increasing function in both arguments, for $y, y^\prime < R$ and $y + y^\prime < R$. Second, for all $y + y^\prime < R$, $h_1(y) \le h_3(y^\prime)$ and $h_2(y) \le h_3(y)$, while $h_3(y), h_3(y^\prime) < 1$, so that
\begin{equation}\label{eq:derivative_hyp_ball_Xi_bounds}
	2 e^{-(R - y - y^\prime)}h_1(y) h_2(y) \le \Xi(y,y^\prime) \le 2 e^{-(R - y - y^\prime)}.
\end{equation}
In particular, since $R-y$ is an increrasing function of $n$ uniformly on $0 < y < (1-\varepsilon)R$, there exists a $0 < \delta < 1$ such that $1/2 \le \Xi(y,y^\prime) < 2$ for all $y + y^\prime < R$ and $(1-\delta)(R-y) < y^\prime < R$ and $n$ large enough.
